\message{ !name(APSA_Paper_2012.tex)}% EBMA paper for Political Analysis: As Accepted for Publication.



%\documentclass[pdftex,12pt,fullpage,oneside,endnotes]{amsart}
\documentclass[12pt,fullpage,endnotes]{article}
%\usepackage{apsr}
\usepackage{array,amsmath,psfrag,amssymb,subfigure,tabularx}
\usepackage{hyperref,multicol}
\usepackage{booktabs}
\usepackage[usenames]{color}
\usepackage{datetime}
\usepackage{dcolumn}
\usepackage{wrapfig}
\usepackage{setspace}
\usepackage{url}
\usepackage[english]{babel}
\usepackage{times}
\usepackage{multirow}
\usepackage[pdftex]{graphicx}
\usepackage{epstopdf}
\usepackage{lscape}
\usepackage{array}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{endnotes}
\usepackage[top=3cm, bottom=3cm, left=2.3cm, right=2.3cm]{geometry} 

\newcommand{\note}[1]{\footnote{ #1 \vspace{4 mm}}}

\usepackage{natbib}
\bibpunct{(}{)}{;}{a}{}{,}
\bibdata{Bibliography_EBMA}
%\bibliographystyle{chicago}

\newboolean{blind}
\setboolean{blind}{false}


\title{Say yes to the guess: \\Ensemble forecasting with sparse data\thanks{This work was supported by the Information Processing Technology Office of the
    Defense Advanced Research Projects Agency through a
    holding grant is to the Lockheed Martin Corporation [FA8650-07-C-7749].}}
\author{
Jacob M. Montgomery\\
	Department of Political Science\\
	Washington University in St. Louis\\
	Campus Box 1063, One Brookings Drive\\
	St. Louis, MO, USA, 63130-4899 
	\and
Florian M. Hollenbach  \\
	Department of Political Science\\
	Duke University\\
	Perkins Hall 326 Box 90204\\
	Durham, NC, USA, 27707-4330
	\and
Michael D. Ward\\
	Department of Political Science\\
	Duke University\\
	Perkins Hall 326 Box 90204\\
	Durham, NC, USA, 27707-4330\\
	corresponding author: michael.d.ward@duke.edu
} 




\date{\today}


\begin{document}

\message{ !name(APSA_Paper_2012.tex) !offset(-3) }


\maketitle
\thispagestyle{empty}
\clearpage
\pagestyle{myheadings}
\markright{Montgomery, Hollenbach, \& Ward\hfill Ensemble BMA\hfill}
\newpage
%\singlespacing

\thispagestyle{empty}


\begin{abstract}
%\begin{doublespace}
 Bla bla blub
 %  \end{doublespace}
\end{abstract}

%\doublespacing
%\newpage


\setcounter{page}{1}

\section{Introduction}

Although accurate prediction of future events is not the primary goal
for most social sciences, recent years have witnessed spreading of
systematic forecasting from more traditional topics (e.g., GDP growth
and unemployment) to many new domains (e.g., elections and mass
killings) .  Several factors have motivated this increase.  To begin
with, testing systematic predictions about future events against
observed outcomes is generally seen as the most stringent validity
check of statistical and theoretical models.  In addition, forecasting
of important political, economic, and social events is of great
interest to policymakers and the general public who are generally less
interested testing theories of the world than correctly anticipating
and altering the future.

With the proliferation of forecasting efforts, however, comes a need
for sensible methods to aggregate and utilize the various scholarly
efforts.  One attractive solutions to this problem is to combine
various prediction models and create an ensemble forecast.  Combining
forecasts reduces reliance on any single data source or methodology,
but also allows for the incorporation of more information than any one
model is likely to include in isolation.  Across subject domains,
ensemble predictions are usually more accurate than any individual
component model. Second, they are signiﬁcantly less likely to make
dramatically incorrect predictions \citep{Bates:1969, Armstrong:2001,
  Raftery:2005}.

The idea of ensemble learning itself has a long history in the machine
learning and nonparametric statistics community. The most thorough
treatment is found in \citet{Hastie:2009}. A wide range of statistical
approaches including neural nets, bagging, random forests, additive
regression trees, and boosting and more may be properly considered
ensemble approaches.  

One ensemble method advocated recently for forecasting is ensemble
Bayesian model averaging (EBMA). This methods was first proposed by
\citet{Raftery:2005} and recently forwarded as a useful method for the
social sciences by \citet{Montgomery:2012a}. In essence, EBMA creates a
finite mixture model that creates a kind of weighted average of
forecasts.  EBMA mixture models seek to collate the good parts of
existing forecasting models while avoiding over-fitting to past
observations or over-estimating our certainty about the future.  The
hope is for greater accuracy as both the knowledge and implied
uncertainty of a variety of approaches are integrated into a combined
predictive probability distribution.

However, there are several challenges for creating ensemble
predictions for many social science applications.  To begin with
amount and quality of data for calibrating ensembles is far from
ideal.  EBMA was first developed for use for weather forecasting
ensembles where measurement of outcomes is fairly precise, and data is
relatively abundant.  Predicting, for instance, water surface
temperatures in 200 locations across five days provides 1,000
observations by which model weights can be calibrated.  Meanwhile, forecasting
quarterly GDP growth in the United States for five years only provides
20.

It is now common for people to make predictions





The first constraint is the paucity of the data.  Relative to weather
and other stuff, there just aren't a lot of observations, mostly
stemming from the fact that for any given time period there is exactly
one outcome.

A second is the large number of forecasting efforts.  For example, the
economic forecasting survey has like a billion experts.

A final issue is the inconsistency with which forecasts are
issued. For any given time window, there are many missing foreacsts.
Moreover, it is not random with larger degrees of misssingness
existing further back in time.

As an example, Table 1 represents nearly the entire population of
legitimate forecasting efforts.  We see all three of these things
going on at the same time.


\begin{table}[ht]
\caption{Pre-election forecasts of the percent of the two-party vote going to the incumbent party in U.S. Presidential elections}
\footnotesize
\begin{center}
\begin{tabular}{rlrrrrrrrrr}
  \toprule
  & F & A & C & H & LBRT & L & Hol & EW & Cuz \\ 
  \midrule
  1992 & 55.7 & 46.3 & 49.7 & 48.9 & 47.3 &  &  &  &  \\ 
  1996 & 49.5 & 57.0 & 55.5 & 53.5 & 53.3 &  & 57.2 & 55.6 &  \\ 
  2000 & 50.8 & 53.2 & 52.8 & 54.8 & 55.4 & 60.3 & 60.3 & 55.2 &  \\ 
  2004 & 57.5 & 53.7 & 52.8 & 53.2 & 49.9 & 57.6 & 55.8 & 52.9 & 51.1 \\ 
  2008 & 48.1 & 45.7 & 52.7 & 48.5 & 43.4 & 41.8 & 44.3 & 47.8 & 48.1 \\ 
  \bottomrule

\end{tabular}
\end{center}
Forecasts were published prior to each election by \textbf{F}air, \textbf{A}bramowitz, \textbf{C}ampbell, \textbf{H}ibbs, \textbf{L}ewis-\textbf{B}eck and \textbf{R}ice (1992), Lewis-Beck and \textbf{T}ien  (1996-2008),   \textbf{L}ockerbie, \textbf{Hol}brook, \textbf{E}rikson and \textbf{W}lezien and \textbf{Cuz}an.  Data taken from CITE SILVER POST HERE.
\end{table}


We adjust the basic EBMA model by better handling missing data, 

\section{Notation and EBMA model} 

Assume a quantity of interest to forecast, $\mathbf{y}^{t^*}$, in some
future period $t^\ast \in T^\ast$.  Further assume that we have extant
forecasts for events $\mathbf{y}^t$ for some past period $t \in T$
that were generated from $K$ forecasting models or teams, $M_1, M_2,
\ldots, M_K$, which have prior probability $M_k\sim \pi(M_k)$. The PDF
for $\mathbf{y}^t$ is $p(\mathbf{y}^t|M_k)$, the predictive PDF for
the quantity of interest is $p(\mathbf{y}^{t^*}|M_k$), the conditional
probability for each model is $p(M_k|\mathbf{y}^t) =
p(\mathbf{y}^t|M_k)\pi(M_k)/\underset{k=1}{\overset{K}{\sum}}p(\mathbf{y}^t|M_k)\pi(M_k)$
and the and the marginal predictive PDF is $p(\mathbf{y}^{t^*}) =
\underset{k=1}{\overset{K}{\sum}}
p(\mathbf{y}^{t^*}|M_k)p(M_k|\mathbf{y}^{t})$.  This can be viewed as
the weighted average of the component PDFs where the weights are
determined by each model's performance within the already-observed
period $T$.



\subsection{Dynamic ensemble forecasting}

The EBMA procedure assumes the prior construction of multiple
forecasting models or heuristics in some training period $T^{\prime}$.
The goal is to estimate the parameters for the ensemble prediction
model using $\mathbf{f}^{t}_k$ for some period $T$.  t is then
possible to generate true ensemble forecasts ($\mathbf{f}_k^{t^\ast}$)
for observations in the test period $t^\ast \in T^*$.  Past
applications have typically statistically post-process the predictions
for out-of-sample bias reduction and treat these re-calibrated
predictions as a component model.  However, in the presence of sparse
data we treat these raw predictions as a component model in the steps
below.  Additional discussion of this is below.

As a running example, let us assume that we have $K$ forecasting
efforts for modeling insurgencies in a set of countries $S$ ongoing
throughout the training ($T^{\prime}$) validation ($T$) and test
($T^\ast$) periods.  We will associate each component forecast with a
component PDF, $g_k(\mathbf{y}|\mathbf{f}_k^{s|t, t^\ast})$, which may
be the original prediction from the forecast model or the
bias-corrected forecast.


The EBMA PDF is then a finite mixture of the $K$ component PDFs,
denoted
\begin{equation}
\small
\label{BMA-eq}
p(\mathbf{y}|\mathbf{f}_1^{s|t}, \ldots, \mathbf{f}_K^{s|t})=\overset{K}{\underset{k=1}{\sum}} w_k
g_k(\mathbf{y}|\mathbf{f}_k^{s|t}),
\end{equation}
\noindent The $w_k$'s $\in [0,1]$ are model probabilities and
$\sum_{k=1}^Kw_k=1$.  Roughly speaking, they are associated with each
component model's predictive performance in the validation period
controlling for the degree to which they offer unique insight (i.e.,
a model's predictions are distinct from those of other component models).
We provide additional discussion about the model weights in the
election forecasting example below.  Details for parameter estimation
are provided in Appendix A.  The ensemble PDF for an insurgency in the
test period $t^\ast$ in country $s$ is then:
\begin{equation}
\label{BMA-eq2}
\small
p(y|f_{1}^{s|t^\ast}, \ldots,
f_{K}^{s|t^\ast})=\overset{K}{\underset{k=1}{\sum}} w_k
g_k(y|f_{k}^{s|t^*}).
\end{equation}


\citet{Raftery:2005} propose approximating the conditional PDF as a
normal distribution centered at a linear transformation of the
individual forecast, $g_k(\mathbf{y}|\mathbf{f}_k^{s|t}) = N(a_{k0} +
a_{k1}\mathbf{f}_k^{s|t}, \sigma^2)$.  Prior applications have found
that this adjustment of the component models' forecasts reduces
over-fitting and improves the performance of the final ensemble
forecasting model \citep{Raftery:2005}.\note{Our adjustments to the
  basic EBMA method for application to dichotomous outcomes, as well
  as details of parameter estimation, are shown in Appendix A.}  Using
\eqref{BMA-eq} and \eqref{BMA-eq2} above, the EBMA PDF is then
\begin{equation} \small
p(\mathbf{y}|\mathbf{f}_1^{s|t}, \ldots, \mathbf{f}_K^{s|t}) = \overset{K}{\underset{k=1}{\sum}} w_k N(a_{k0} +
a_{k1}\mathbf{f}_k^{s|t}, \sigma^2),
\end{equation}
\noindent and the predictive distribution for some observation $y$ is 
\begin{equation} \small
p(y|f_1^{s|t^\ast}, \ldots, f_K^{s|t^\ast}) = \overset{K}{\underset{k=1}{\sum}} w_k N(a_{k0} +
a_{k1}f_k^{s|t^\ast}, \sigma^2)
 \end{equation}
 \noindent Thus, the predictive PDF is a mixture of $K$ normal
 distributions each of whose mean is determined by the component
 prediction ($f_k^{s|t^\ast}$) and whose ``height'' (i.e, the total
 area under the curve for component $k$) is determined by the model
 weight $w_k$.

\subsection{Estimation}

\subsection{Missing forecast components}

%\section{Window selection} Jacob thinks this should be a short subsection eventually.  But not for APSA.

\section{Small sample adjustment}

%\section{Simulations} Jacob think we will want just a bit of this.  But not for APSA.

\section{Applications}
\subsection{Predicting unemployment}
\subsection{Predicting presidential elections}


\section{Discussion} 

% \newpage
% \appendix


% \section*{Appendix}

%%Bib 
\singlespacing
\bibliographystyle{apsr}
\bibliography{Bibliography_EBMA}


\end{document}
\bye

\message{ !name(APSA_Paper_2012.tex) !offset(-328) }
